---
title: "Employee Turnover"
output:
  word_document: default
  html_document: default
date: "2022-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(mlr)
library(survival)
library(survminer)
library(tidyverse)
library(MASS)
# library(pec)
# library(survAUC)
library(dplyr)
# library(reshape2)
library(ggplot2)
library(lubridate)
```

## 1 Introduction

We chose to analyze the Employee Turnover dataset *insert footnote*. This Employer Turnover dataset contains 1129 observations with 17 covariates about how long it took until the employer quit the job. We are using the column stag as our time variable, which represents failure and censor time in units months *verified*. In the event column, 1 is for an employee quitting, 0 is for a censor. 


The covariates in our model include the character variables: gender, age, industry, profession, traffic, coach, head_gender, greywage, and way. Additional covariates include float variables ranging from 1-10: extraversion, independence, selfcontrol, anxiety, novator. *change variables potentially?, see:* "The features self-control, anxiety, independ, and novator were renamed conscientiousness, neuroticism, agreeableness and openness to experience, based on the Big Five Personality Model." (L.J.C.H. Hamers, 2021) 


Our project will aim to predict the probability of an employee turning over or leaving versus not leaving based on the events and variables in the dataset. *One key question asked by the study.* In reference to the dataset, "the goal is to predict individual risks of quitting of specific applicants." (E. Babushkin, 2017). *Add citation to information.* 


*Questions asked in analysis:* influence of personality traits and other covariates on turnover. (relevant plots, will include later)

```{r}
# code chunk, insert plots here 
```

*Explanatory summary of final model*: 



"To answer the posed research questions, this study made use of the Employee Turnover
dataset which contains information on personality traits, demographics, industry, and coaching. It is a real dataset shared by E. Babushkin (2017), validated by research of Baginska (2018). Edward Babushkin, an HR analyst from Moscow, performed himself survival analysis to predict the risk of particular employees’ quitting (Babushkin, 2017). The data stems from a call-center company with various branches across Russia. Turnover is a key issue within this industry." (L.J.C.H. Hamers, 2021) 

http://arno.uvt.nl/show.cgi?fid=157305 


```{r, echo=FALSE}
turnover <- read_csv("turnover.csv")
turnover <- turnover %>% mutate(stagSurv = Surv(stag, event),
                                # difference in change of factoring?
                                gender = as.factor(ifelse(gender=="m","Male","Female")),
                                coach = as.factor(coach),
                                greywage = as.factor(greywage),
                                as.factor(ifelse(head_gender == "m","Male","Female")),
                                industry = as.factor(industry),
                                profession = as.factor(profession),
                                traffic = as.factor(traffic),
                                way = as.factor(way))
#turnover$stag*487*12/15
# changed yellow to green because hard to see in plot
roygbiv <- c("red","orange","green","cyan","blue","slateblue2","violet","black")
fourcolor <- function(total, front){
  return(c(rep("white", front), "black", "red", "blue", "green", rep("white", total-front-4)))}
```

# Forward AIC and BIC Models

Since there are no more than one event occurring per observation, we are going to look at the normal cox ph model.
```{r}
CPH.all <- coxph(stagSurv~gender+age+industry+traffic+coach+head_gender+greywage+way+
                           extraversion+independ+selfcontrol+anxiety+novator,turnover)
# CPH.all
``` 
(gender refactor?) Looking at the p-values, we remove the least relevant observations because of on their large p-values. 

Running through the stepAIC function, we get the final model of the following:
```{r}
CPH.AICfit <- coxph(formula = stagSurv ~ age+industry+profession+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```
Now running through the forward stepwise AIC function, we can determine that age, industry, profession, traffic, greywage, way(method of transportation), as well as personality score of self control and anxiety creates a significant difference in the survival rate.


# Proportional Hazards Assumptions (Tetsuto's Version)

We looked through log-log plots for quantitative covariates that have few levels, and run through coxzph test for continuous covariates. For industry and profession, we will go reference the coxph function to group together levels into 3 groups (absolute z value less than 1.282, 1.282 to 2.576, and beyond 2.576)
```{r}
indG1 <- c("Agriculture","Banks","Building","Consult")
indG2 <- c("etc","HoReCa","manufacture","Mining","Pharma","PowerGeneration","State","transport")
profG1 <- c("BusinessDevelopment","Consult","etc","Finan\xf1e","HR","IT","Law","Sales","Teaching")
profG2 <- c("Commercial","Engineer","Marketing","PR")
turnover <- turnover %>% mutate(ind=ifelse(industry %in% indG1,"G1",
                                           ifelse(industry %in% indG2,"G2","G3")),
                                prof=ifelse(profession %in% profG1,"G1",
                                           ifelse(profession %in% profG2,"G2","G3")))
CPH.Groupfit <- coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```


We looked through log-log plots for quantitative covarites, and run through coxzph test for continuous covariates.
```{r}
par(mfrow=c(2,3))
# Tetsuto's version 

# ggsurvplot(survfit(stagSurv ~ gender, turnover), fun = "cloglog", censor.size=0.4, size = 0.4, xlab="Time in Months", title="Stratified Gender") # , col = roygbiv) # include gender? refactoring influence?

# might not include Gender if not statistically significant
plot(survfit(stagSurv ~ ind, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Gender", col = roygbiv)
plot(survfit(stagSurv ~ ind, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Industry", col = roygbiv)
plot(survfit(stagSurv ~ prof, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Profession", col = roygbiv)
plot(survfit(stagSurv ~ traffic, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Traffic", col = roygbiv)
plot(survfit(stagSurv ~ greywage, turnover), fun = "cloglog", xlab="Time in Months", main="Stratified Greywage", col = roygbiv)
plot(survfit(stagSurv ~ way, turnover), fun = "cloglog", xlab="Time in Months", main="Stratified Way", col = roygbiv)

CZPH.fit <- cox.zph(CPH.Groupfit)
CZPH.fit
```
All the graphs gave mostly parallel lines and cox zph test gives a global p-value of greater than 0.05.

# Proportional Hazards Assumptions (Ezra's Version)
```{r}
par(mfrow=c(2,3))
colorlines <- c("blue", "pink", "red", "hotpink", "maroon", "orange", "yellow", 
                "green", "turquoise", "darkgreen", "violet", "lavender", "black", 
                "brown", "magenta", "gold")
# Gender Log(-Log) Plot
gender.KM <- survfit(stagSurv ~ as.factor(gender), data = turnover)
plot(gender.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Gender Log(-Log) curves")
legend(x="topleft", title="Legend", legend=c("Male", "Female"), fill=c("blue", "pink"))
# Industry Log(-Log) Plot
industry.KM <- survfit(stagSurv ~ as.factor(industry), data = turnover)
plot(industry.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Industry Log(-Log) curves")
# Profession Log(-Log) Plot
profession.KM <- survfit(stagSurv ~ as.factor(profession), data = turnover)
plot(profession.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Profession Log(-Log) curves")
# Traffic Log(-Log) Plot
traffic.KM <- survfit(stagSurv ~ as.factor(traffic), data = turnover)
plot(traffic.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Traffic Log(-Log) curves")
# Greywage Log(-Log) Plot
greywage.KM <- survfit(stagSurv ~ as.factor(greywage), data = turnover)
plot(greywage.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Greywage Log(-Log) curves")
legend(x="topleft", title="Legend", legend=c("grey", "white"), fill=c("blue", "pink"))
# Way Log(-Log) Plot
way.KM <- survfit(stagSurv ~ as.factor(way), data = turnover)
plot(way.KM, fun = "cloglog", col = colorlines, xlab = "time",
     ylab = "Log(-Log) survival", main = "Way Log(-Log) curves")
legend(x="topleft", title="Legend", legend=c("bus", "car", "foot"), fill=c("blue", "pink", "red"))
```

\newpage

## Goodness-of-Fit Approach to Testing the PH Assumption

With this approach we utilize a test statistic and p-value in order to assess the validity of the PH assumption for any given predictor. This is a more concrete and accurate method to test this assumption in comparison to the graphical approach. For this study we will use the `cox.zph()` function to give us the p-values of the Global and Individual Shoenfeld Tests, which we will also represent with a plot using the `ggcoxzph()` function.


```{r, echo=FALSE}
# Cox PH Model: Gender
coxph.gender <- coxph(stagSurv~gender, data=turnover)
summary(coxph.gender)$coefficients
# anova(coxph.gender)
testzph.gender <- cox.zph(coxph.gender)
ggcoxzph(testzph.gender)
# Cox PH Model: Industry
coxph.industry <- coxph(stagSurv~industry, data=turnover)
summary(coxph.industry)$coefficients
# anova(coxph.industry)
testzph.industry <- cox.zph(coxph.industry)
ggcoxzph(testzph.industry)$coefficients
# Cox PH Model: Profession
coxph.profession <- coxph(stagSurv~profession, data=turnover)
summary(coxph.profession)$coefficients
# anova(coxph.profession)
testzph.profession <- cox.zph(coxph.profession)
ggcoxzph(testzph.profession)$coefficients[]
# Cox PH Model: Traffic
coxph.traffic <- coxph(stagSurv~traffic, data=turnover)
summary(coxph.traffic)$coefficients
# anova(coxph.traffic)
testzph.traffic <- cox.zph(coxph.traffic)
ggcoxzph(testzph.traffic)
# Cox PH Model: Grey Wage
coxph.greywage <- coxph(stagSurv~greywage, data=turnover)
summary(coxph.greywage)$coefficients
# anova(coxph.greywage)
testzph.greywage <- cox.zph(coxph.greywage)
ggcoxzph(testzph.greywage)
# Cox PH Model: Way
coxph.way <- coxph(stagSurv~way, data=turnover)
summary(coxph.way)$coefficients
# anova(coxph.way)
testzph.way <- cox.zph(coxph.way)
ggcoxzph(testzph.way)
```


*will probably need help with interpretation and tying into conclusion section* 

## 4 Conclusions
• Hazard Ratios
• 95% confidence intervals for these hazard ratios

*from template*: "In your introduction, you should have set a scientific question that you are answering. You need to report your conclusion to that question. This will often be something along the lines of “the chemotherapy treatment showed a statistically significant decrease in the hazard rate for the time until re-infection.” I also want for you to report the hazard ratio for the relevant factors to your research question. In addition, I want 95% confidence intervals for these hazard ratios. These parameter values and intervals can be read directly from the summary of the coxph model, but I want to see these numbers pulled out into the narrative of the report." 

https://stats.stackexchange.com/questions/379469/how-to-interpret-hazard-ratios-of-cox-output

### Conclusion starts here

```{r, echo=FALSE}
summary(CPH.Groupfit)
```

```{r}
# remove exp(coef)
summary(CPH.Groupfit)$conf.int[,-1]
```
*relevancy of liklihood ratio test??*  

```{r}
# adjusted GroupFit (looking at the p-values)
CPH.Groupfit <- coxph(formula = stagSurv ~ gender+age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```

# Additional Models

Now we are going to make adjustments to personality scores to determine if any of the personality scores affect the time in anyway. Change the score to be 0 if less than 5, and subtract score by 5 otherwise.*reasoning for not including all personality metrics?*

do we want to follow procedures like in lecture or attempt exponential model effect
```{r, include=FALSE}
turnover <- turnover %>% mutate(ex5 = ifelse(extraversion <= 5, 0, extraversion-5),
                                id5 = ifelse(independ <= 5, 0, independ-5),
                                sc5 = ifelse(selfcontrol <= 5, 0, selfcontrol-5),
                                ax5 = ifelse(anxiety <= 5, 0, anxiety-5),
                                nv5 = ifelse(novator <= 5, 0, novator-5))
CPH.max<-coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety+
      ex5+id5+sc5+ax5+nv5, data = turnover)
summary(CPH.max)$coefficients[18:22,]
```

```{r, incluude=FALSE}
# reasoning for removing sc5 and ax5f from model? 
# summary(coxph(formula = stagSurv ~ age+ind+prof+traffic+
#     greywage+way+selfcontrol+anxiety+
#       ex5+id5+nv5, data = turnover))$coefficients
stepAIC(coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety+
      ex5+id5+nv5, data = turnover))
CPH.cutoff5 <- coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+
    way+selfcontrol+anxiety+ex5, data = turnover)
cox.zph(CPH.cutoff5)
```


We will also try the same for cutoff at 3 and 7
```{r}
turnover <- turnover %>% mutate(ex3 = ifelse(extraversion <= 3, 0, extraversion-3),
                                id3 = ifelse(independ <= 3, 0, independ-3),
                                sc3 = ifelse(selfcontrol <= 3, 0, selfcontrol-3),
                                ax3 = ifelse(anxiety <= 3, 0, anxiety-3),
                                nv3 = ifelse(novator <= 3, 0, novator-3),
                                ex7 = ifelse(extraversion <= 7, 0, extraversion-7),
                                id7 = ifelse(independ <= 7, 0, independ-7),
                                sc7 = ifelse(selfcontrol <= 7, 0, selfcontrol-7),
                                ax7 = ifelse(anxiety <= 7, 0, anxiety-7),
                                nv7 = ifelse(novator <= 7, 0, novator-7))
CPH.max <-coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+
                  extraversion+independ+selfcontrol+anxiety+novator+
                  ex3+id3+sc3+ax3+nv3+
                  ex5+id5+sc5+ax5+nv5+
                  ex7+id7+sc7+ax7+nv7, data = turnover)
# CPH.cutoff.fit <- stepAIC(CPH.max)
# cox.zph(CPH.cutoff.fit)
```


```{r}
CPH.max <-coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety+
      ex3 + id3 + sc3 + ax3 + nv3, data = turnover)
# CPH.cutoff.fit <- stepAIC(CPH.max)
# cox.zph(CPH.cutoff.fit)

CPH.cutoff5 <- coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+selfcontrol+anxiety+ex5, data = turnover)
# cox.zph(CPH.cutoff5)
```


ex3,ex7,and ax7 has higher p-value than ex5 and ax5 respectively, so our final model will exclude them
```{r}
coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+ex5+ax5+sc7,data = turnover)
```


### 1.1 Citations

Babushkin, E. (2017). Employee turnover: how to predict individual risks of quitting.
Retrieved from https://edwvb.blogspot.com/2017/10/employee-turnover-how-to-predict-individual-risks-of-quitting.html

Hamers, L.J.C.H. (2021). Retrieved from: http://arno.uvt.nl/show.cgi?fid=157305 

Kleinbaum, David G., and Mitchel Klein. 2012. Survival Analysis, a Self-Learning Text. New York: Springer.

Potential References??:
https://stats.stackexchange.com/questions/379469/how-to-interpret-hazard-ratios-of-cox-output, https://stats.stackexchange.com/questions/560975/how-to-interpret-schoenfield-residual-plot 

