---
title: "NYPD_Shooting_Incident"
output:
  pdf_document: default
  html_document: default
date: "2022-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(mlr)
library(survival)
library(survminer)
library(pec)
library(survAUC)
library(MASS)
library(dplyr)
library(reshape2)
library(ggplot2)
library(readr)
```

```{r}
turnover <- read_csv("turnover.csv")
turnover <- turnover %>% mutate(stagSurv = Surv(stag, event),
                                gender = as.factor(gender),
                                coach = as.factor(coach),
                                greywage = as.factor(greywage),
                                head_gender = as.factor(head_gender),
                                industry = as.factor(industry),
                                profession = as.factor(profession),
                                traffic = as.factor(traffic),
                                way = as.factor(way))
#turnover$stag*487*12/15
color <- c("red","orange","yellow","green","blue","slateblue2","violet","black")
fourcolor <- function(total, front){
  return(c(rep("white", front), "black", "red", "blue", "green", rep("white", total-front-4)))
}
```
Since there are no more than one event occuring per observation, we are going to look at the normal cox ph model.
```{r}
CPH.all <- coxph(stagSurv~gender+age+industry+traffic+coach+head_gender+greywage+way+
                           extraversion+independ+selfcontrol+anxiety+novator,turnover)
```

Running through the stepAIC function, we get the final model of the following:
```{r}
CPH.AICfit <- coxph(formula = stagSurv ~ age+industry+profession+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```
Now running through the forward stepwise AIC function, we can determine that age, industry, profession, traffic, greywage, way(method of transportation), as well as personality score of self control and anxiety creates a significant difference in the survival rate.

We looked through log-log plots for quantitative covariates that have few levels, and run through coxzph test for continuous covariates. For industry and profession, we will go reference the coxph function to group together levels into 3 groups (absolute z value less than 1.282, 1.282 to 2.576, and beyond 2.576)
```{r}
indG1 <- c("Agriculture","Banks","Building","Consult")
indG2 <- c("etc","HoReCa","manufacture","Mining","Pharma","PowerGeneration","State","transport")
profG1 <- c("BusinessDevelopment","Consult","etc","Finan\xf1e","HR","IT","Law","Sales","Teaching")
profG2 <- c("Commercial","Engineer","Marketing","PR")
turnover <- turnover %>% mutate(ind=ifelse(industry %in% indG1,"G1",
                                           ifelse(industry %in% indG2,"G2","G3")),
                                prof=ifelse(profession %in% profG1,"G1",
                                           ifelse(profession %in% profG2,"G2","G3")))
CPH.Groupfit <- coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```

```{r}
plot(survfit(stagSurv ~ ind, turnover), fun = "cloglog", col = color)
plot(survfit(stagSurv ~ prof, turnover), fun = "cloglog", col = color)
plot(survfit(stagSurv ~ traffic, turnover), fun = "cloglog", col = color)
plot(survfit(stagSurv ~ greywage, turnover), fun = "cloglog", col = color)
plot(survfit(stagSurv ~ way, turnover), fun = "cloglog", col = color)
CZPH.fit <- cox.zph(CPH.Groupfit)
CZPH.fit
```
All the graphs gave mostly parallel lines. However, the industry lines is small, so we will look ath the Schoenfeld residuals plot.
```{r}
ggcoxzph(CZPH.fit, var = "ind")
```


We will look at the stratification  to 
```{r}
# stepAIC(CPH.strataAge)
CPH.sAge.fit <- coxph(formula = stagSurv ~ strata(age)+industry+traffic+
                        greywage+way+selfcontrol, data = turnover)
CZPH.sAge <- cox.zph(CPH.sAge.fit)
CZPH.sAge
```

```{r}
ggcoxzph(CZPH.sAge, var = c("industry","traffic"))
```
greywage, way, and selfcontrol are not significant, so we can assume proportional hazards, but industry, traffic, and global test reject the significant test, so we will test for stratification

```{r}
CPH.strata2 <- coxph(formula = stagSurv ~ strata(age)+strata(industry)+strata(traffic)+
                        greywage+way+selfcontrol, data = turnover)
CZPH.strata2 <- cox.zph(CPH.strata2)
CZPH.strata2
ggcoxzph()
```


We will run through BIC as well to confirm.
```{r}
CPH.null <- coxph(formula = stagSurv ~ 1,
    data = turnover)
# step(CPH.null, scope = list(lower=CPH.null,upper=CPH.all),
#      direction="both", criterion = "BIC", k=log(length(turnover)))
CPH.BICfit <- coxph(formula = stagSurv ~ age+greywage+way+extraversion,
                    data = turnover)
```
BIC gave a different model for us to use, but for now we wil be following through with the AIC model for now.



grey wage tends to quit earlier than white wage
bus > car > foot, not sure about the reason why

"Agriculture"     "Banks"           "Building"        "Consult"         "etc"            
"HoReCa"          "IT"              "manufacture"     "Mining"          "Pharma"         
"PowerGeneration" "RealEstate"      "Retail"          "State"           "Telecom"        
"transport" 

```{r}
```

plot(survfit(turnover$stagSurv ~ 1), main = "Survival Curve of turnover time",
     xlab="Time (Months?)", ylab="Probability of Survival",
     font.main=4, font.lab=4, font.sub=4, col = "darkorchid3")


Look at the result and ph assumptions
Hazard ratio, conclusion, adjustments to the model
Looking at the personality pieces, that hazard ratios (95% CI)

Beyond that... try something like
extrovert > 5
0     if ext. <= 5, 
ext-5 if ext. > 5






