---
title: "NYPD_Shooting_Incident"
output:
  pdf_document: default
  html_document: default
date: "2022-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(mlr)
library(survival)
library(survminer)
library(pec)
library(survAUC)
library(MASS)
library(dplyr)
library(reshape2)
library(ggplot2)
library(readr)
```

```{r}
turnover <- read_csv("turnover.csv")
turnover <- turnover %>% mutate(stagSurv = Surv(stag, event),
                                gender = as.factor(gender),
                                coach = as.factor(coach),
                                greywage = as.factor(greywage),
                                head_gender = as.factor(head_gender),
                                industry = as.factor(industry),
                                profession = as.factor(profession),
                                traffic = as.factor(traffic),
                                way = as.factor(way))
#turnover$stag*487*12/15
color <- c("red","orange","lightgreen","green","blue","slateblue2","violet","black")
fourcolor <- function(total, front){
  return(c(rep("white", front), "black", "red", "blue", "green", rep("white", total-front-4)))
}
```
Since there are no more than one event occuring per observation, we are going to look at the normal cox ph model.
```{r}
CPH.all <- coxph(stagSurv~gender+age+industry+traffic+coach+head_gender+greywage+way+
                           extraversion+independ+selfcontrol+anxiety+novator,turnover)
```

Running through the stepAIC function, we get the final model of the following:
```{r}
CPH.AICfit <- coxph(formula = stagSurv ~ age+industry+profession+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
```
Now running through the forward stepwise AIC function, we can determine that age, industry, profession, traffic, greywage, way(method of transportation), as well as personality score of self control and anxiety creates a significant difference in the survival rate.

We looked through log-log plots for quantitative covariates that have few levels, and run through coxzph test for continuous covariates. For industry and profession, we will go reference the coxph function to group together levels into 3 groups (absolute z value less than 1.282, 1.282 to 2.576, and beyond 2.576)
```{r}
indG1 <- c("Agriculture","Banks","Building","Consult")
indG2 <- c("etc","HoReCa","manufacture","Mining","Pharma","PowerGeneration","State","transport")
profG1 <- c("BusinessDevelopment","Consult","etc","Finan\xf1e","HR","IT","Law","Sales","Teaching")
profG2 <- c("Commercial","Engineer","Marketing","PR")
turnover <- turnover %>% mutate(ind=ifelse(industry %in% indG1,"G1",
                                           ifelse(industry %in% indG2,"G2","G3")),
                                prof=ifelse(profession %in% profG1,"G1",
                                           ifelse(profession %in% profG2,"G2","G3")))
CPH.Groupfit <- coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety, data = turnover)
ggcoxzph(cox.zph(coxph(stagSurv~profession)))
```

```{r}
par(mfrow=c(2,3))
plot(survfit(stagSurv ~ ind, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Industry", col = color)
plot(survfit(stagSurv ~ prof, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Profession", col = color)
plot(survfit(stagSurv ~ traffic, turnover), fun="cloglog", xlab="Time in Months", main="Stratified Traffic", col = color)
plot(survfit(stagSurv ~ greywage, turnover), fun = "cloglog", xlab="Time in Months", main="Stratified Greywage", col = color)
plot(survfit(stagSurv ~ way, turnover), fun = "cloglog", xlab="Time in Months", main="Stratified Way", col = color)

CZPH.fit <- cox.zph(CPH.Groupfit)
CZPH.fit
```
All the graphs gave mostly parallel lines and cox zph test gives a global p-value of greater than 0.05.

#Conclusion starts here
```{r}
as.matrix(round(exp(-CPH.Groupfit$coefficients),3))
print("")
round(exp(-confint(CPH.Groupfit)),3)
```

#Additional Models

Now we are going to make adjustments to personality scores to determine if any of the personality scores affect the time in anyway. Change the score to be 0 if less than 5, and subtract score by 5 otherwise.
```{r}
turnover <- turnover %>% mutate(ex3 = ifelse(extraversion <= 3, 0, extraversion-3),
                                id3 = ifelse(independ <= 3, 0, independ-3),
                                sc3 = ifelse(selfcontrol <= 3, 0, selfcontrol-3),
                                ax3 = ifelse(anxiety <= 3, 0, anxiety-3),
                                nv3 = ifelse(novator <= 3, 0, novator-3),
                                ex5 = ifelse(extraversion <= 5, 0, extraversion-5),
                                id5 = ifelse(independ <= 5, 0, independ-5),
                                sc5 = ifelse(selfcontrol <= 5, 0, selfcontrol-5),
                                ax5 = ifelse(anxiety <= 5, 0, anxiety-5),
                                nv5 = ifelse(novator <= 5, 0, novator-5),
                                ex7 = ifelse(extraversion <= 7, 0, extraversion-7),
                                id7 = ifelse(independ <= 7, 0, independ-7),
                                sc7 = ifelse(selfcontrol <= 7, 0, selfcontrol-7),
                                ax7 = ifelse(anxiety <= 7, 0, anxiety-7),
                                nv7 = ifelse(novator <= 7, 0, novator-7))
coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety+
      ex5+id5+nv5, data = turnover)
stepAIC(coxph(formula = stagSurv ~ age+ind+prof+traffic+
    greywage+way+selfcontrol+anxiety+
      ex5+id5+nv5, data = turnover))
CPH.cutoff5 <- coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+
    way+selfcontrol+anxiety+ex5, data = turnover)
cox.zph(CPH.cutoff5)
```

We will also try the same for cutoff at 3 and 7
```{r}
turnover <- turnover %>% mutate()
CPH.max <-coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+
                  extraversion+independ+selfcontrol+anxiety+novator+
                  ex3+id3+sc3+ax3+nv3+
                  ex5+id5+sc5+ax5+nv5+
                  ex7+id7+sc7+ax7+nv7, data = turnover)
CPH.cutoff.fit <- stepAIC(CPH.max)
stepAIC(coxph(formula = stagSurv ~ selfcontrol*anxiety, data = turnover))
stepAIC(coxph(formula = stagSurv ~ selfcontrol*anxiety + selfcontrol + anxiety, data = turnover))
```

```{r}
CPH.3 <- coxph(formula = stagSurv ~ gender+age+industry+traffic+coach+head_gender+greywage+way+
                  ex3+id3+sc3+ax3+nv3, data = turnover)
CPH.5 <- coxph(formula = stagSurv ~ gender+age+industry+traffic+coach+head_gender+greywage+way+
                  ex5+id5+sc5+ax5+nv5, data = turnover)
CPH.7 <- coxph(formula = stagSurv ~ gender+age+industry+traffic+coach+head_gender+greywage+way+
                  ex7+id7+sc7+ax7+nv7, data = turnover)
CPH.cutoff3 <- stepAIC(CPH.3)
CPH.cutoff5 <- stepAIC(CPH.5)
CPH.cutoff7 <- stepAIC(CPH.7)
cox.zph(CPH.cutoff.fit)
summary(coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+ex5+ax5+sc7,data = turnover))$conf.int[16:18,]
```

ex3,ex7,and ax7 has higher p-value than ex5 and ax5 respectively, so our final model will exclude them
```{r}
coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+
        ex3+ex5+ax5+ex7+sc7+ax7,data = turnover)
coxph(formula = stagSurv ~ age+ind+prof+traffic+greywage+way+
        ex5+ax5+sc7,data = turnover)
```



