---
title: "NYPD_Shooting_Incident"
output: html_document
date: '2022-11-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(mlr)
library(survival)
library(pec)
library(survAUC)
library(dplyr)
library(reshape2)
library(ggplot2)
```

```{r}
turnover <- read_csv("turnover.csv")
turnover <- turnover %>% mutate(stagSurv = Surv(stag, event),
                                gender = as.factor(ifelse(gender == "m","Male","Female")),
                                coach = as.factor(coach),
                                greywage = as.factor(greywage),
                                head_gender = as.factor(ifelse(head_gender == "m","Male","Female")),
                                industry = as.factor(industry),
                                profession = as.factor(profession),
                                traffic = as.factor(traffic))
#turnover$stag*487*12/15
CPH.personality <- coxph(stagSurv~extraversion+independ+selfcontrol+anxiety+novator,turnover)
stepAIC(CPH.personality)
CPH.fit <- coxph(formula = stagSurv ~ selfcontrol + anxiety, data = turnover)
coxph(formula = stagSurv ~ selfcontrol*anxiety, data = turnover)
stepAIC(coxph(formula = stagSurv ~ selfcontrol*anxiety+selfcontrol+anxiety, data = turnover))
```
interaction between covariates are not significant to be included in the model.

```{r}
BIC(CPH.personality)
BIC(CPH.fit)
```
Since CPH.fit gives smaller BIC score than CPH.personality, we can confirm it is better fit model

```{r}
```

plot(survfit(turnover$stagSurv ~ 1), main = "Survival Curve of turnover time",
     xlab="Time (Months?)", ylab="Probability of Survival",
     font.main=4, font.lab=4, font.sub=4, col = "darkorchid3")


Look at the result and ph assumptions
Hazard ratio, conclusion, adjustments to the model
Looking at the personality pieces, that hazard ratios (95% CI)

Beyond that... try something like
extrovert > 5
0     if ext. <= 5, 
ext-5 if ext. > 5






